Fail Faster

### 2025-03-16 paper/paper.tex
minor

  % in strict functionallanguages like OCaml and Scal
To answer this, we propose a two-pronged approach for designing generator libraries 
that preserves their expressiveness while eliminating unnecessary overhead.
First, we describe a method of optimizing
generator libraries using \textit{staged metaprogramming}, or ``staging'', which has 
long been used to eliminating overhead 
from idiomatic functional DSLs \todo{CITE!!!}.
Using staging, it is possible to eliminate the overhead of many common 
generator abstractions. For example, in unstaged generator libraries, each call to monadic bind constructs and immediately executes a closure at runtime.
Meanwhile, a staged approach can resolve the bind at compile time, enabling the compiler to inline and eliminate these closures.
Secondly, we demonstrate the
impact of different choices of \rand on generator performance 
by conducting a controlled comparison. Together, these insights provide a
recipe for improving even state-of-the-art generator libraries.

@@

To evaluate \name's effectiveness, we
implement a series of workloads in each generator library, and show that 
\name runs faster and allocates less than \\\bq. Further, we show that this performance benefit
translates to bug-finding ability by running these workloads in Etna, a platform that 
simulates real-world PBT usage by injecting bugs into a system and measuring how quickly various 
generators detect them.
Finally, to demonstrate the generality of our technique, we implement \name in Scala
and benchmark the same generators in the Scala port of the library---thereby highlighting 
its portability
to other strict functional languages and reinforcing its potential for widespread adoption.

In summary, we establish that monadic PBT generator DSLs
have a significant performance overhead across languages, and present a cross-language technique
for eliminating that overhead while preserving their idiomatic abstractions.
We make the following contributions:
\begin{enumerate}
    \item We identify two key sources of inefficiencies in PBT generator libraries, which can significantly impact performance: 
    namely, \rand choice, and abstraction overhead. We describe a general procedure to address the latter based on staged metaprogramming. \secref{section:motiv}
    \item We implement a version of \bq, \name, that showcases our optimization technique while maintaining strict program equality. \secref{section:4}
    \item We evaluate \name by demonstrating the improved performance of generators constructed
    using our library in a controlled comparison, as well as its impact on bug-finding ability. \secref{section:eval}
\end{enumerate}

>>>>

  % in strict functionallanguages like OCaml and Scal
To answer this, we propose a two-pronged approach for designing generator libraries 
that preserves their expressiveness while eliminating unnecessary overhead.
First, we describe a method of optimizing
generator libraries using \textit{staged metaprogramming}, or ``staging'', which has 
long been used for opetimizing idiomatic functional DSLs \todo{CITE!!!}.
We extend this technique to generators, completely eliminating the overhead of many common generator abstractions. For example, in unstaged generator libraries, each call to monadic bind constructs and immediately executes a closure at runtime.
Meanwhile, a staged approach can resolve the bind at compile time, enabling the compiler to inline and eliminate these closures.
Secondly, we demonstrate the
impact of different choices of \rand on generator performance 
by conducting a controlled comparison. Together, these insights provide a
recipe for improving even state-of-the-art generator libraries.

@@

To evaluate \name's effectiveness, we
implement a series of workloads in each generator library, and show that 
\name runs faster and allocates less than \\\bq. Further, we show that this performance benefit
translates to bug-finding ability by running these workloads in Etna, a platform that 
simulates real-world PBT usage by injecting bugs into a system and measuring how quickly various 
generators detect them.
Finally, to demonstrate the generality of our technique, we implement \name in Scala
and benchmark the same generators in the Scala port of the library---thereby highlighting 
its portability
to other strict functional languages.

In summary, we show that monadic PBT generator DSLs incur significant performance costs across languages, and present a general technique for improving their efficiency while preserving their idiomatic abstractions.
We make the following contributions:
\begin{enumerate}
    \item We identify two key sources of inefficiencies in PBT generator libraries, which can significantly impact performance: 
    namely, \rand choice, and abstraction overhead. \secref{section:motiv}
    \item We apply staged metaprogramming to the novel domain of generators. \secref{section:}
    \item We implement a version of \bq, \name, that showcases our optimization technique while maintaining strict program equality. \secref{section:4}
    \item We evaluate \name by demonstrating the improved performance of generators constructed
    using our library in a controlled comparison, as well as its impact on bug-finding ability. \secref{section:eval}
\end{enumerate}
